Some examples in extended logic counting classes

### 2015-10-09 tex/classes.tex
22 sep
8 sep

\section{Extended Horn counting hierarchies}

We define syntactically the classes $\EiE$ and $\EiU$ as following. First we define Horn clauses.
\begin{eqnarray*}
PL &::=& X_i(\x),\, i\in\N,\\
NL &::=& \neg X_i(\x),\, i\in\N \ \mid \ \exists x\, NL,\\
NC &::=& NL \ \mid \ \alpha, \alpha \mbox{ is an {\sc FO}-formula over } \L \ \mid \ (NC \vee NC),\\
HC &::=& NC \ \mid \ (NC \vee PL),
\end{eqnarray*}
where $\x$ is a tuple with the corresponding number of variables. Now we define the classes' syntaxis inductively.
\begin{enumerate}
	\item $\EzeroE$:
	\begin{eqnarray*}
	E_0 &::=& HC \ \mid \ E_0 \wedge E_0.
	\end{eqnarray*}
	\item $\EzeroU$:
	\begin{eqnarray*}
	U_0 &::=& E_0.
	\end{eqnarray*}
	\item $\EipE$:
	\begin{eqnarray*}
	E_{i+1} &::=& U_i \ \mid \ \exists x \, E_{i+1}.
	\end{eqnarray*}
	\item $\EipU$:
	\begin{eqnarray*}
		U_{i+1} &::=& E_i \ \mid \ \forall x \, U_{i+1}.
	\end{eqnarray*}
\end{enumerate}
A function $f$ is in $\EiE$ (resp. $\EiU$) if there is an $\L$-formula $\varphi$ defined by the grammar $E_i$ (resp. $U_i$) such that $f = f_{\varphi}$.

@@

	We will now prove that there is a polynomial time algorithm that decides $L_{f}$. This is equivalent to decide, given $\A= \langle A, \S\A, \leq^\A \rangle\in\Truc$, whether there exist assignments $\P,\z$ such that $\A\models\varphi(\P,\z)$. This is the same as $\A\models\exists\X\exists\x\,\varphi(\X,\x)$. Since $\varphi(\X,\x)$ is defined by the grammar $E_2$, there exists $\psi(\X,\x,\y,\u)$ such that $\varphi(\X,\x) = \exists\x\exists\y\forall\u\exists\v\,\psi(\X,\x,\y,\u,\v)$, where every variable in $\v$ appears on a negated second-order literal $\neg X_i$, with $i\in\N$.

	Given $\A\in\Truc$ we generate an equivalent $\L$-formula $\theta(\X,\x,\y)$ with a series of operations. First, we replace every instance of $\v$ by a disjunction of all $r$-tuples in $A^r$, where $r$ is the number of variables in $\v$. The remainder is still a {\sc Horn}-clause. Second, we replace every instance of $\u$ by a conjunction of all $s$-tulples in $A^s$, where $s$ is the number or variables in $\u$. The remainder is still a conjunction of {\sc Horn}-clauses.

>>>>

\section{The Extended Horn Counting Hierarchy}

We define syntactically the classes $\EiE$ and $\EiU$ as follows. First we define extended Horn clauses.
\begin{eqnarray*}
PL &::=& X_i(\x),\, i\in\N,\\
NL &::=& \neg X_i(\x),\, i\in\N \ \mid \ \exists x\, NL,\\
NC &::=& NL \ \mid \ \alpha, \alpha \mbox{ is an {\sc FO}-formula over } \L \ \mid \ (NC \vee NC),\\
HC &::=& NC \ \mid \ (NC \vee PL),
\end{eqnarray*}
where $\x$ is a tuple with the corresponding number of variables. Now we define the syntax of the classes inductively.
\begin{enumerate}
	\item $\EzeroE$:
	\begin{eqnarray*}
	E_0 &::=& HC \ \mid \ E_0 \wedge E_0.
	\end{eqnarray*}
	\item $\EzeroU$:
	\begin{eqnarray*}
	U_0 &::=& E_0.
	\end{eqnarray*}
	\item $\EipE$:
	\begin{eqnarray*}
	E_{i+1} &::=& U_i \ \mid \ \exists x \, E_{i+1}.
	\end{eqnarray*}
	\item $\EipU$:
	\begin{eqnarray*}
		U_{i+1} &::=& E_i \ \mid \ \forall x \, U_{i+1}.
	\end{eqnarray*}
\end{enumerate}
A function $f$ is in $\EiE$ (resp. $\EiU$) if there is an $\L$-formula $\varphi$ defined by the grammar $E_i$ (resp. $U_i$) such that $f = f_{\varphi}$.

@@

	We will now prove that there is a polynomial time algorithm that decides $L_{f}$. This is equivalent to decide, given $\A= \langle A, \S^\A, \leq^\A \rangle\in\Truc$, whether there exist assignments $\P,\z$ such that $\A\models\varphi(\P,\z)$. This is the same as $\A\models\exists\X\exists\x\,\varphi(\X,\x)$. Since $\varphi(\X,\x)$ is defined by the grammar $E_2$, there exists $\psi(\X,\x,\y,\u)$ such that $\varphi(\X,\x) = \exists\x\exists\y\forall\u\exists\v\,\psi(\X,\x,\y,\u,\v)$, where every variable in $\v$ appears on a negated second-order literal $\neg X_i$, with $i\in\N$.

	Given $\A\in\Truc$ we generate an equivalent $\L$-formula $\theta(\X,\x,\y)$ with a series of operations. First, we replace every instance of $\v$ by a disjunction of all $r$-tuples in $A^r$, where $r$ is the number of variables in $\v$. The result is still a Horn clause. Second, we replace every instance of $\u$ by a conjunction of all $s$-tulples in $A^s$, where $s$ is the number or variables in $\u$. The result is still a conjunction of Horn clauses.

### 2015-10-09 tex/examples.tex
8 sep

\subsection{{\sc \#3-HORN-SAT}}
We will now show that {\sc \#3-HORN-SAT} $\in\EoneU$.\\

Let $\L = \{P_1,P_2,P_3,N_1,N_2,N_3,\leq\}$, and $\A = \langle A, P_1^\A,P_2^\A,P_3^\A,N_1^\A,N_2^\A,N_3^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc 3-HORN-SAT} formula $\Psi$, where $A$ is the set of variables mentioned in $\Psi$, and the relations in $\L$ are interpreted as follows:
\begin{eqnarray*}
P_1^\A &=& \{a\mid a \mbox{ appears as a conjunct in }\Psi\},\\
P_2^\A &=& \{(a_1,a_2)\mid (a_1\vee \neg a_2) \mbox{ appears as a conjunct in }\Psi\},\\
P_3^\A &=& \{(a_1,a_2,a_3)\mid (a_1\vee \neg a_2\vee \neg a_3) \mbox{ appears as a conjunct in }\Psi\},\\
N_1^\A &=& \{a\mid (\neg a) \mbox{ appears as a conjunct in }\Psi\},\\
N_2^\A &=& \{(a_1,a_2)\mid (\neg a_1\vee \neg a_2) \mbox{ appears as a conjunct in }\Psi\},\\
N_3^\A &=& \{(a_1,a_2,a_3)\mid (\neg a_1\vee \neg a_2\vee \neg a_3) \mbox{ appears as a conjunct in }\Psi\}.
\end{eqnarray*}

@@

We can easily reorder $\varphi(T)$ as $\forall x \forall y \forall z\,\theta(x,y,z,T)$ where $\theta(x,y,z,T)$ is an {\sc FO}-extended quantifier-free $\L$-formula such that any second order variable only appears in a Horn clause. It is easily seen that each relation $T$ over $\A$ such that $\A\models\forall x \forall y \forall z\,\theta(x,y,z,T)$ represents a valid assignment to the variables in $\Psi$, and therefore, $f_{\forall x \forall y \forall z\,\theta(x,y,z,T)}$ represents {\sc \#3-HORN-SAT}.

@@

\subsection{{\sc \#HORN-SAT}}
We will now show that {\sc \#HORN-SAT} $\in\EoneU$.\\

Let $\L = \{P,N,F,\leq\}$, and $\A = \langle A, P^\A, N^\A, F^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc HORN-SAT} formula $\Psi$, where $A$, $P$ and $N$ are defined as in the previous case, and $F$ is defined as
\[
	F = \{ c \mid
		\mbox{all literals in $c$ are negative}
	\}
\]

Now let $T$ be a unary second-order variable, and
\begin{align*}
	\psi(T) &=& \forall c\,[ F(c) \to \exists \ell'(N(c,\ell') \wedge \neg T(\ell'))]
	\,\wedge
	 \forall c \forall \ell\, [\neg F(c)\to(P(c,\ell)\wedge\forall\ell'(N(c,\ell')\to T(\ell')))\to T(\ell)] \\
	 &\equiv& \forall c\,[ \neg F(c) \vee \exists \ell'(N(c,\ell') \wedge \neg T(\ell'))]
	 \,\wedge
	 \forall c \forall \ell\, [ F(c) \vee \neg P(c,\ell) \vee \exists\ell'(N(c,\ell') \wedge \neg T(\ell'))) \vee T(\ell)]
\end{align*}
Let $\neg X(c,\ell')$ be defined as $N(c,\ell')\wedge \neg T(\ell)$. That is,
\begin{eqnarray*}
	&& \forall c \forall \ell'\, [(N(c,\ell') \wedge \neg T(\ell')) \leftrightarrow \neg X(c,\ell')]\\
	&\equiv& \forall c \forall \ell'\,[(\neg N(c,\ell')\vee T(\ell')\vee \neg X(c,\ell'))\wedge (X(c,\ell')\vee \neg T(\ell'))\wedge (X(c,\ell')\vee N(c,\ell'))].
\end{eqnarray*}
Finally, let
\begin{multline*}
	\varphi(T,X) = \forall c \forall \ell'\,[(\neg N(c,\ell')\vee T(\ell')\vee \neg X(c,\ell'))\wedge (X(c,\ell')\vee \neg T(\ell'))\wedge (X(c,\ell')\vee N(c,\ell'))]\,\wedge \\
	\forall c\,[ \neg F(c) \vee \exists \ell'\neg X(c,\ell'))]
	\wedge
	\forall c \forall \ell\, [ F(c) \vee \neg P(c,\ell) \vee \exists\ell'\neg X(c,\ell') \vee T(\ell)].
\end{multline*}
Let $\theta(c,\ell,T,X)$ be such that $\varphi'(T,X) = \forall c \forall \ell\,\theta(c,\ell,T,X)$. Note that $\theta(c,\ell,T,X)$ is an {\sc FO}-extended quantifier-free $\L$-formula such that any second order variable only appears in a Horn clause. First, note that each relation $T$ over $\A$ such that $\A\models \psi(T)$ represents a valid assignment to the variables in $\Psi$, and that in $\theta(c,\ell,T,X)$, $X$ is completely defined by the assignment to $T$. Therefore $f_{\psi(T)}(\A') = f_{\varphi'(T,X)}(\A')$ for each $\A'\in\Truc$, from which we conclude that $f_{\forall c \forall \ell\,\theta(c,\ell,T,X)}$ represents {\sc \#HORN-SAT}.

\subsection{{\sc \#Disj-HORN-SAT}}
We define the decision problem
\[
	\textsc{Disj-HORN-SAT} = \{\Phi \mid \Phi \mbox{ is a disjunction of Horn formulas and $\Phi$ is satisfiable}\},
\]
and the counting problem {\sc \#Disj-HORN-SAT} as a function that counts all satisfying assignments to a formula $\Phi$ which is a disjunction of Horn formulas. We will now show that {\sc \#Disj-HORN-SAT} $\in\EtwoE$.

Let $\L = \{P,N,F,B,\leq\}$, and $\A = \langle A, P^\A, N^\A, F^\A, B^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc Disj-HORN-SAT} formula $\Psi$ with $n$ variables, $m$ clauses in total, and $k$ Horn formulas, where $A = \{1,\ldots,n+m+k\}$, $P$, $N$ and $F$ are defined as in the previous case, and $B$ is defined as:
\[
	B = \{ (c,f) \mid 
		\mbox{the clause $c$ is in the Horn formula $f$}
	\}.
\]
Now let $T$ be a unary second order variable, and
\[
	\psi(T) = \exists f\forall c\,(B(c,f)\to \forall\ell\,\theta(c,\ell,T,X)),
\]
where $\theta$ is defined as in the previous case.

>>>>

\subsection{{\sc \#3-Horn-Sat}}
We will now show that {\sc \#3-Horn-Sat} $\in\EoneU$.\\

Let $\L = \{P_1,P_2,P_3,N_1,N_2,N_3,\leq\}$, and $\A = \langle A, P_1^\A,P_2^\A,P_3^\A,N_1^\A,N_2^\A,N_3^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc 3-Horn-Sat} formula $\Psi$, where $A$ is the set of variables mentioned in $\Psi$, and the relations in $\L$ are interpreted as follows:
\begin{eqnarray*}
P_1^\A &=& \{a\mid a \mbox{ appears as a conjunct in }\Psi\},\\
P_2^\A &=& \{(a_1,a_2)\mid (a_1\vee \neg a_2) \mbox{ appears as a conjunct in }\Psi\},\\
P_3^\A &=& \{(a_1,a_2,a_3)\mid (a_1\vee \neg a_2\vee \neg a_3) \mbox{ appears as a conjunct in }\Psi\},\\
N_1^\A &=& \{a\mid (\neg a) \mbox{ appears as a conjunct in }\Psi\},\\
N_2^\A &=& \{(a_1,a_2)\mid (\neg a_1\vee \neg a_2) \mbox{ appears as a conjunct in }\Psi\},\\
N_3^\A &=& \{(a_1,a_2,a_3)\mid (\neg a_1\vee \neg a_2\vee \neg a_3) \mbox{ appears as a conjunct in }\Psi\}.
\end{eqnarray*}

@@

We can easily reorder $\varphi(T)$ as $\forall x \forall y \forall z\,\theta(x,y,z,T)$ where $\theta(x,y,z,T)$ is an {\sc FO}-extended quantifier-free $\L$-formula such that any second order variable only appears in a Horn clause. It is easily seen that each relation $T$ over $\A$ such that $\A\models\forall x \forall y \forall z\,\theta(x,y,z,T)$ represents a valid assignment to the variables in $\Psi$, and therefore, $f_{\forall x \forall y \forall z\,\theta(x,y,z,T)}$ represents {\sc \#3-Horn-Sat}.

@@

\subsection{{\sc \#Horn-Sat}}
We will now show that {\sc \#Horn-Sat} $\in\EoneU$.\\

Let $\L = \{P,N,F,\leq\}$, and $\A = \langle A, P^\A, N^\A, F^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc Horn-Sat} formula $\Psi$, where $A$, $P$ and $N$ are defined as in the previous case, and $F$ is defined as
\[
	F = \{ c \mid
		\mbox{all literals in $c$ are negative}
	\}
\]

Now let $T$ be a unary second-order variable, and
\begin{align*}
	\psi(T) &=& \forall c\,[ F(c) \to \exists \ell'(N(c,\ell') \wedge \neg T(\ell'))]
	\,\wedge
	 \forall c \forall \ell\, [\neg F(c)\to(P(c,\ell)\wedge\forall\ell'(N(c,\ell')\to T(\ell')))\to T(\ell)] \\
	 &\equiv& \forall c\,[ \neg F(c) \vee \exists \ell'(N(c,\ell') \wedge \neg T(\ell'))]
	 \,\wedge
	 \forall c \forall \ell\, [ F(c) \vee \neg P(c,\ell) \vee \exists\ell'(N(c,\ell') \wedge \neg T(\ell'))) \vee T(\ell)]
\end{align*}
Let $\neg X(c,\ell')$ be defined as $N(c,\ell')\wedge \neg T(\ell)$. That is,
\begin{eqnarray*}
	&& \forall c \forall \ell'\, [(N(c,\ell') \wedge \neg T(\ell')) \leftrightarrow \neg X(c,\ell')]\\
	&\equiv& \forall c \forall \ell'\,[(\neg N(c,\ell')\vee T(\ell')\vee \neg X(c,\ell'))\wedge (X(c,\ell')\vee \neg T(\ell'))\wedge (X(c,\ell')\vee N(c,\ell'))].
\end{eqnarray*}
Finally, let
\begin{multline*}
	\varphi(T,X) = \forall c \forall \ell'\,[(\neg N(c,\ell')\vee T(\ell')\vee \neg X(c,\ell'))\wedge (X(c,\ell')\vee \neg T(\ell'))\wedge (X(c,\ell')\vee N(c,\ell'))]\,\wedge \\
	\forall c\,[ \neg F(c) \vee \exists \ell'\neg X(c,\ell'))]
	\wedge
	\forall c \forall \ell\, [ F(c) \vee \neg P(c,\ell) \vee \exists\ell'\neg X(c,\ell') \vee T(\ell)].
\end{multline*}
Let $\theta(c,\ell,T,X)$ be such that $\varphi'(T,X) = \forall c \forall \ell\,\theta(c,\ell,T,X)$. Note that $\theta(c,\ell,T,X)$ is an {\sc FO}-extended quantifier-free $\L$-formula such that any second order variable only appears in a Horn clause. First, note that each relation $T$ over $\A$ such that $\A\models \psi(T)$ represents a valid assignment to the variables in $\Psi$, and that in $\theta(c,\ell,T,X)$, $X$ is completely defined by the assignment to $T$. Therefore $f_{\psi(T)}(\A') = f_{\varphi'(T,X)}(\A')$ for each $\A'\in\Truc$, from which we conclude that $f_{\forall c \forall \ell\,\theta(c,\ell,T,X)}$ represents {\sc \#Horn-Sat}.

\subsection{{\sc \#Disj-Horn-Sat}}
We define the decision problem
\[
	\textsc{Disj-Horn-Sat} = \{\Phi \mid \Phi \mbox{ is a disjunction of Horn formulas and $\Phi$ is satisfiable}\},
\]
and the counting problem {\sc \#Disj-Horn-Sat} as a function that counts all satisfying assignments to a formula $\Phi$ which is a disjunction of Horn formulas. We will now show that {\sc \#Disj-Horn-Sat} $\in\EtwoE$.

Let $\L = \{P,N,F,B,\leq\}$, and $\A = \langle A, P^\A, N^\A, F^\A, B^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc Disj-Horn-Sat} formula $\Psi$ with $n$ variables, $m$ clauses in total, and $k$ Horn formulas, where $A = \{1,\ldots,n+m+k\}$, $P$, $N$ and $F$ are defined as in the previous case, and $B$ is defined as:
\[
	B = \{ (c,f) \mid 
		\mbox{the clause $c$ is in the Horn formula $f$}
	\}.
\]
Now let $T$ be a unary second order variable, and
\[
	\psi(T) = \exists f\forall c\,(B(c,f)\to \forall\ell\,\theta(c,\ell,T,X)),
\]
where $\theta$ is defined as in the previous case.

### 2015-10-09 tex/s-one-fo.tex
8 sep

\begin{theo}
	{\sc \#DNF} is hard for $\E1$ under parsimonious reductions. 
\end{theo}
\begin{proof}
	Let $f$ be an arbitrary function in $\E1$ and let $f_{\textsc{DNF}}$ be the function that characterizes {\sc \#DNF}. We will now show a function $h:\Truc\to L(P)$ such that for every $\A\in\Truc$, it holds that $f(\A) = f_{\textsc{DNF}}(h(\A))$, and which can be computed in polynomial time. Let $\psi$ be such that $f = f_{\psi}$. We separate the proof in three cases:
	\begin{enumerate}
		\item All free variables in $\psi$ are of first order. Let $\psi = \exists x\,\varphi(\x,\z)$ where $\x = (x_1,\ldots,x_c)$ and $\z = (z_1,\ldots,z_d)$. Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$, we generate a propositional formula $\Phi = f(\A)$. First we notice that this case is trivial since we can calculate this function in polynomial time, by checking for each $\e\in A^d$ if $\A\models\exists\x\,\varphi(\x,\e)$. Let $n$ be the number of such assignments to $\z$. Let $\p$ be an $n$-tuple of propositional variables where $\p = (p_1,\ldots,p_n)$. The DNF formula we return is
		\[
		\Phi = \bigvee_{i = 1}^n \neg p_1 \wedge \cdots \wedge \neg p_{i-1} \wedge p_i \wedge \neg p_{i+1} \wedge \cdots \wedge \neg p_n
		\]
		which has exactly $n$ possible assignments.
		\item There is only second-order free variables in $\psi$. Let $\psi = \exists x\,\varphi(\x,\X)$, where $\x = (x_1,\ldots,x_c)$ and $\X = (X_1,\ldots,X_r)$. Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$, we generate a DNF-formula $\Phi = f(\A)$ as follows:

@@

		Note that for each assignment $\P$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\P)$ there is a corresponding assignment $\sigma$ to $\p$ where $\sigma(p^{\ell}_{\a}) = 1$ if and only if $\a\in P_{\ell}$. We can conclude that $f(\A) = f_{\textsc{DNF}}(\Phi)$.
		\item Let $\psi = \exists x\,\varphi(\x,\X,\z)$, where $\x = (x_1,\ldots,x_c)$, $\X = (X_1,\ldots,X_r)$ and $\z = (z_1,\ldots,z_d)$. Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$ we generate a DNF-formula $\Phi = f(\A)$ as follows:

@@

		We generate a propositional variable $p^{\e}_{\ell,\b}$ for each $\b\in \bigcup_{j = 1}^{c+d} A^j$, for each $\e\in A^d$, and for each $\ell\in\{1,\ldots,r\}$. We also generate a $t_{\a}$ for each $\a$ Let $\p$ be the tuple of all such variables. Also, let $\rho_{\ell}$ be such that $\rho_{\ell}(\x,\z) = \y_{\ell}$

		For each $\e\in A^d$, we follow the following procedure. For each $\varphi_i(\x,\X)$, we generate the set $\Gamma^{\e}_i = \{\a\in A^c\mid \A\models\alpha(\a,\e)\}$, and so, the DNF formula we generate is
		\[
		\Phi^{\e}_i = \bigvee_{\a\in\Gamma_i} p^{\e}_{s_1,\rho_{s_1}(\a,\e)} \wedge \cdots \wedge p^{\e}_{s_m,\rho_{s_m}(\a,\e)} \wedge \neg p^{\e}_{t_1,\rho_{t_1}(\a,\e)} \wedge \cdots \wedge \neg p^{\e}_{t_n,\rho_{t_n}(\a,\e)} \wedge t_{\e},
		\]
		where $\beta_{\e}$ is defined as

	\end{enumerate}
\end{proof}

>>>>

\begin{theo}
	{\sc \#DNF} is hard for $\E1$ under parsimonious reductions. 
\end{theo}
\begin{proof}
	Let $f$ be an arbitrary function in $\E1$ and let $f_{\textsc{DNF}}$ be the function that characterizes {\sc \#DNF}. We will now show a function $h:\Truc\to L(P)$ such that for every $\A\in\Truc$, it holds that $f(\A) = f_{\textsc{DNF}}(h(\A))$, and which can be computed in polynomial time. Let $\psi$ be such that $f = f_{\psi}$. We separate the proof in three cases:
	\begin{enumerate}
		\item All free variables in $\psi$ are of first order. Let $\psi = \exists \x\,\varphi(\x,\z)$ where $\x = (x_1,\ldots,x_c)$ and $\z = (z_1,\ldots,z_d)$. Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$, we generate a propositional formula $\Phi = h(\A)$. First we notice that this case is trivial since we can calculate this function in polynomial time, by checking for each $\e\in A^d$ if $\A\models\exists\x\,\varphi(\x,\e)$. Let $n$ be the number of such assignments to $\z$. Let $\p$ be an $n$-tuple of propositional variables where $\p = (p_1,\ldots,p_n)$. The DNF formula we return is
		\[
		\Phi = \bigvee_{i = 1}^n \neg p_1 \wedge \cdots \wedge \neg p_{i-1} \wedge p_i \wedge \neg p_{i+1} \wedge \cdots \wedge \neg p_n
		\]
		which has exactly $n$ possible assignments.

		\item There is only second-order free variables in $\psi$. Let $\psi = \exists \x\,\varphi(\x,\X)$, where $\x = (x_1,\ldots,x_c)$ and $\X = (X_1,\ldots,X_r)$. Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$, we generate a DNF-formula $\Phi = h(\A)$ as follows:

@@

		Note that for each assignment $\P$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\P)$ there is a corresponding assignment $\sigma$ to $\p$ where $\sigma(p^{\ell}_{\a}) = 1$ if and only if $\a\in P_{\ell}$. We can conclude that $f(\A) = f_{\textsc{DNF}}(\Phi)$.

		\item Let $\psi = \exists \x\,\varphi(\x,\X,\z)$, where $\x = (x_1,\ldots,x_c)$, $\z = (z_1,\ldots,z_d)$, $\X = (X_1,\ldots,X_r)$ and each $X_{\ell}$ has arity $c_{\ell}$. Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$ we generate a DNF-formula $\Phi = h(\A)$ as follows:

@@

		We generate a propositional variable $p^{\e}_{\ell,\b}$ for each $\e\in A^d$, for each $\ell\in\{1,\ldots,r\}$, and for each $\b\in A^{c_{\ell}}$. Let $\p$ be the tuple of all such variables. We also generate a set $T = \{t_1,\ldots,t_{L}\}$ where $L = \lceil \log\vert A^d\vert \rceil$, and an arbitrary inyective function $\beta:A^d \to \{0,1\}^L$ where $\beta(\e) = (b^{\e}_1,\ldots,b^{\e}_L)$ for each $\e\in A^d$. Also, let $\rho_{\ell}$ be such that $\rho_{\ell}(\x,\z) = \y_{\ell}$.

		For each $\e\in A^d$, we follow the following procedure. For each $\varphi_i(\x,\X)$, we consider the set $\Gamma^{\e}_i = \{\a\in A^c\mid \A\models\alpha(\a,\e)\}$, and so, the DNF formula we generate is
		\[
		\Phi^{\e}_i = \bigvee_{\a\in\Gamma_i} p^{\e}_{s_1,\rho_{s_1}(\a,\e)} \wedge \cdots \wedge p^{\e}_{s_m,\rho_{s_m}(\a,\e)} \wedge \neg p^{\e}_{t_1,\rho_{t_1}(\a,\e)} \wedge \cdots \wedge \neg p^{\e}_{t_n,\rho_{t_n}(\a,\e)} \wedge \gamma_{\e} \wedge \theta_{\e}
		\]
		where $\gamma_{\e}$ is defined as
		\[
			\gamma_{\e} = \bigwedge_{\e'\in A^d\setminus\{\e\}} \bigwedge_{\b\in \bigcup_{j = 1}^{c+d} A^j} \bigwedge_{\ell = 1}^r p_{\ell,\b}^{\e'},
		\]
		and $\theta_{\e}$ is defined as
		\[
			\theta_{\e} = \bigwedge_{i\in[1,L]:\,b^{\e}_i = 0} \neg t_i\  \wedge \bigwedge_{i\in[1,L]:\,b^{\e}_i = 1} t_i.
		\]
		Finally, we return $f(\A) = \Phi = \bigvee_{i = 1}^k \bigvee_{\e\in A^d} \Phi_i^{\e}$.

		We will now show a function $g:\{\langle\P,\e\rangle \mid \A \models \exists\x\,\varphi(\x,\P,\e)\} \to \{\sigma \mid \sigma(\Phi) = 1\}$ and then prove that it is a bijection. Let $g$ be such that for each assignment $\langle \P,\e \rangle$, it holds that $g(\langle \P,\e \rangle) = \sigma$ where $\sigma$ is constructed as follows:

		For each $t_i\in T$, let $\sigma(t_i) = b^{\e}_i$. For each $p_{\ell,\b}^{\e}$, where $\ell\in\{1,\ldots,r\}$ and $\b\in A^{c_{\ell}}$, let $\sigma(p_{\ell,\b}^{\e}) = 1$ if and only if $\b \in P_{\ell}$. For each $p_{\ell,\b}^{\e'}$, where $\e'\in A^d\setminus\{\e\}$,  $\ell\in\{1,\ldots,r\}$ and $\b\in A^{c_{\ell}}$, let $\sigma(p_{\ell,\b}^{\e'}) = 1$. Let $i$ be such that $\A\models \exists\x\,\varphi_i(\x,\P,\e)$. It is simple to see that from the way we constructed $\Phi_i^{\e}$, it holds that $\sigma(\Phi_i^{\e}) = 1$. Also, note that every different assignment $\langle \P',\e' \rangle$, it holds that $\sigma(\langle \P',\e' \rangle) \neq \sigma(\langle \P,\e \rangle)$. Therefore, $g$ is an injective function. Now, let $\sigma$ be an arbitrary assignment such that $\sigma(\Phi) = 1$. Let $i$ and $\e$ be such that $\sigma(\Phi^{\e}_i)$. Note that for every $\e'\in A^d\setminus\{\e\}$, it holds that $\sigma(\theta_{\e'}) = 0$, so there is exactly one possible $\e$ associated to $\sigma$. Consider the assignment $\P$ where $\b\in P_{\ell}$ if and only if $\sigma(p^{\e}_{\ell,\b}) = 1$, for each $\ell$ and $\b \in A^{c_{\ell}}$. Since there is at least one $\a\in\Gamma_i$ for which $\sigma$ satisfies its respective subformula, it holds that $\A\models\varphi_i(\a,\P,\e)$, which implies that $\A\models\exists\x\,\varphi(\x,\P,\e)$. From this we deduce that $g$ is surjective, and therefore a bijection. We conclude that $f(\A) = f_{\textsc{DNF}}(\Phi)$.
	\end{enumerate}
\end{proof}
